---
title: "Loading data with thymusatlastools2"
author: "Eric Kernfeld"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading data with thymusatlastools2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


You can load Maehr Lab data or metadata by using functions from `thymusatlastools2`.

```{r, setup}
knitr::opts_chunk$set(results = "hide")
```

```{r}

our_stuff <- thymusatlastools2::get_metadata()
head(our_stuff[c("Sample_ID", "dge_path")])# better for static PDF of vignette
# View(our_stuff[c("Sample_ID", "dge_path")]) # better if running interactively
Tera_reps <- thymusatlastools2::load_maehrlab_data(c("perturb_2018JAN_hES_tera",
                                                     "perturb_2018JAN_hDE_tera_rep1",
                                                     "perturb_2018JAN_hDE_tera_rep2"))
```

If you check the results, you'll see that the return value is a list of sparse matrices, one per replicate.

```{r}
typeof(Tera_reps)
length(Tera_reps)
names (Tera_reps)
dim   (Tera_reps[["perturb_2018JAN_hDE_tera_rep1"]])
typeof(Tera_reps[["perturb_2018JAN_hDE_tera_rep1"]])
```

There is a utility `dge_merge_list` for merging raw data. It's fairly memory-efficient, and it handles lists of arbitrary length. 

```{r}
for(i in 1:3){
  Tera_reps[[i]] = Tera_reps[[i]][1:100, 1:50]
}
all_TERA_merged <- thymusatlastools2::dge_merge_list(Tera_reps); 
rm(Tera_reps); gc()
typeof(all_TERA_merged)
dim(all_TERA_merged)
```

Once you build a Seurat object, you can use `add_maehrlab_metadata` to integrate more information from the Maehr Lab metadata sheet.

```{r}
all_TERA_seurat = Seurat::CreateSeuratObject(all_TERA_merged, names.delim = "\\|", names.field = 2) 
rm(all_TERA_merged); gc()
all_TERA_seurat = thymusatlastools2::add_maehrlab_metadata(all_TERA_seurat, "source_tissue")
all_TERA_seurat = thymusatlastools2::add_maehrlab_metadata(all_TERA_seurat, "species")
table( all_TERA_seurat@meta.data[["orig.ident"]], 
       all_TERA_seurat@meta.data[["source_tissue"]])
```
